Initial Analysis

### 2019-05-21 Initial_Analysis.Rmd
Initial methods (quite advanced)

#### Library ####
packages <- c(
  "readxl", # Read dataframe
  "dplyr", # Data manipulation
  "tidyr", # Data manipulation
  "ggplot2", #Nice grpahs and spatial analysis
  "cowplot",
  "rgdal",
  "RColorBrewer",
  "knitr",
  "kableExtra",
  "data.table",
  "ggrepel",
  "gridExtra",
  "ggmap",
  "rgeos",
  # "stringr",
  # "stringer",
  "spdep",
  "sf", #Spatial analysis 
  "sp", #Spatial analysis 
  "purrr",#Spatial analysis 
  # "rgdal", #Spatial analysis 
  "tools", #Spatial analysis 
  "png", # For reading plots in chunk codes
  "grid" # For reading plots in chunk codes
)

@@

@@

```{r Spatial_Data, eval=T, echo=T, results='hide'}

@@

# World_EEZ <- readOGR(dsn = path_world,
                     # layer =file_path_sans_ext(fnam_world))

World_EEZs <- st_read(dsn = path_world,
                      layer = file_path_sans_ext(fnam_world))

@@

# DBEM exploited species list 
# exploited_species_list <- fread("/Volumes/DATA/PROJECTION exploited species ENM/exploited_species_list.csv")

# Use thisone for testing 
Selected_Species <- fread("~/Documents/UBC/Oceans_Project/Distribution/By Cell/Cell_Data/List_Species.csv") %>% 
  select(1:2)

@@

### Determine which countries are neighbors

```{r Testing_Neighbors, eval=F, echo=T}
# Determine which countries are neighbors

@@

# Subset data for testing
South_America <-c("Mexico","Argentina","Chile","Peru","Ecuador","Colombia","Venezuela","Brazil")

South_Americasp <- World_EEZ[World_EEZ@data$Name %in% South_America, ]

#Look for comments in previous chunk

# Get the list using the poly2nb function
Neighbor_List <- poly2nb(South_Americasp)

@@

# Get the names of the countries for identifying them
Country_Names <- data_frame(id = row.names(South_Americasp@data),
                            Country_Territory = as.character(South_Americasp@data$Territory1),
                            Neighbor_Territory = Country_Territory)

# Clean up and transform the neighbor matrix
SA_Neighbors <- as.data.frame(Neighbor_Matrix) %>%
  mutate(country = row.names(.)) %>%  # Convert row names to actual column
  gather(neighbor, present, -country) %>%  # Convert to long
  filter(present == 1) %>%  # Only look at cells with a match
  # Add country names
  left_join(select(Country_Names,
                   -Neighbor_Territory), by=c("country" = "id")) %>%
  left_join(select(Country_Names, 
                   -Country_Territory), by=c("neighbor" = "id")) %>% 
  select(4,5) # select final column

head(SA_Neighbors)

# Remove duplicates (e.g. Country A - Country B and Country B - Country A)
### NOTE, I'm going to use the already created one. Load it fiz it and save it with the same name.

SA_Neighbors_Clear <- SA_Neighbors[!duplicated(SA_Neighbors[1:length(SA_Neighbors)]), ]

#### Works beautifully! 
```

```{r Determine_neighbors_Test, eval = F, echo=T}

#### SF package

SAU_Regions <- as.data.frame(unique(World_EEZs$Name))

Test_Countries <-c("Mexico (Atlantic)","Mexico (Pacific)","Guatemala (Caribbean)","Guatemala (Pacific)","Belize")

# Test_sp <- World_EEZ[World_EEZ@data$Name %in% Test_Countries, ]

Test_sf <- filter(World_EEZs, Name %in% Test_Countries)

# ggplot(Test_sf) +
#   geom_sf() # 5 min

# Adds to the shapefile a column with the neighbourds 
Neighbours_sf <- Test_sf %>%
  mutate(Neighbours = map(.x = geometry, 
                          .f = st_intersects, 
                          y = st_geometry(Test_sf))
         ) %>% 
  st_transform(3832) %>% 
  st_simplify(10000)

head(Neighbours_sf)

@@

### Merge Neighbors with DBEM data

```{r Data_Wrangling, eval=T, echo=T, results='hide'}

### Match DBEM list of countries wth Neighbors
Matching_Names <- Neighbours_sf %>% 
  semi_join(EEZIDs_List,
            by ="EEZID") # Only half of them match

head(Matching_Names)

@@

@@

# Varibales needed
# Spp: Species to be abalized, data with presence/absence per gridcell
# Index_EEZ: Reference list of all INDEX that fall within EEZs
# Neighbors: Reference list of all the neighboring EEZs that each country has

# The function
Trans_Spp_Fun <- function(Spp,Index_EEZ,Neighbors){

  Trans_Spp <- Spp %>%
    filter(V1 >= 1) %>% # Filter only presence
    left_join(Index_EEZ,
              by = "INDEX") %>% # Match the species distribution to the Countries EEZs
    group_by(Country_Territory=Name,
             Neighbor_Territory) %>% 
    summarise(N_INDEX = n()) %>% # The number of cells present within each country's EEZ
    mutate(Trans = 1) # Include one transboundary species

  # Result 1. Number of Countries that share the species
  Trans_Spp <- Neighbors %>% 
    left_join(Trans_Spp,
              by = c("Country_Territory","Neighbor_Territory")) %>% 
    mutate_if(is.numeric, funs(replace(., is.na(.),0))) #%>% # Replaces NAs for 0's to sum
    # mutate(Species = ifelse(Trans > 0,paste(Spp[1,2]),NA)) # Includes the species in question

@@

  Countries_with_Specie <- Trans_Spp %>% 
    filter(Trans == 1) %>% 
    group_by(Neighbor_Territory) %>% 
    summarise()

@@

  # print(Countries_with_Specie) # Prints the countries that share such species
  return(Trans_Spp) # A dataframe with each country and neighbouring states sharing the species
} # End function

### Testing Function ####
# WORKS!!!!

# Load Species data
# D_Path <- paste("./Data/123",".csv",sep="")

# Reads file
# Species <- read.csv(D_Path)

# Call the function (Works)
# Function_Test <- Trans_Spp_Fun(Species,DBEM_Neighbors_Data,Neighbors);View(Function_Test)

>>>>

#### Library ####
packages <- c(
  "readxl", # Read dataframe
  "data.table", # Read dataframe (Fast!)
  "dplyr", # Data manipulation
  "tidyr", # Data manipulation
  "ggplot2", #Nice grpahs and spatial analysis
  # "cowplot",
  # "rgdal",
  # "RColorBrewer",
  # "knitr",
  # "kableExtra",
  # "ggrepel",
  # "gridExtra",
  # "ggmap",
  # "rgeos",
  # "stringr",
  # "stringer",
  "spdep", # for poly2nb
  "sf", #Spatial analysis 
  "sp", #Spatial analysis 
  "purrr",#Spatial analysis
  "rgdal", #Spatial analysis
  "tools"#, #Spatial analysis 
  # "png", # For reading plots in chunk codes
  # "grid" # For reading plots in chunk codes
)

@@

### Check Gabs Data ####

# Does all data have the same number of files?

SAU <- length(list.files("/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Distribution_Data/SAU_Distribution/"))
ENM <- length(list.files("/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Distribution_Data/ENM/"))
Occ <- length(list.files("/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Distribution_Data/Occurence/"))

####### Yes they do! ####

@@

```{r Spatial_Data, eval = T, echo = T, results='hide'}

@@

World_EEZs <- readOGR(dsn = path_world,
                      layer =file_path_sans_ext(fnam_world))

# World_EEZs_sf <- st_read(dsn = path_world,
#                       layer = file_path_sans_ext(fnam_world))

@@

#### Estimate the number of grid-cells that each EEZ as

EEZ_Size <- EEZ_CellID %>% 
  group_by(EEZID) %>% 
  summarise(n_grid = n()) %>% 
  left_join(EEZIDs_List,
            by = "EEZID")

@@

## Functions

### Function `EstNeighbors`

This function determines what EEZs are neighbors and returns a data.frame of it

```{r funEstNeighbors, eval = T, echo=T}
# Determine which countries are neighbors

@@

EstNeighbors <- function(Shapefile){

  # Get the list using the poly2nb function
  Neighbor_List <- poly2nb(Shapefile)

@@

  # Get the names of the countries for identifying them
  Country_Names <- data_frame(id = row.names(Shapefile@data),
                              Country_Territory = as.character(Shapefile@data$Name),
                              Neighbor_Territory = Country_Territory,
                              EEZID = Shapefile@data$EEZID
                              )

  # Clean up and transform the neighbor matrix
  Neighbors <- as.data.frame(Neighbor_Matrix) %>%
    mutate(country = row.names(.)) %>%  # Convert row names to actual column
    gather(neighbor, present, -country) %>%  # Convert to long
    filter(present == 1) %>%  # Only look at cells with a match
    # Add country names
    left_join(select(Country_Names,
                     -Neighbor_Territory), by=c("country" = "id")) %>%
    left_join(select(Country_Names, 
                     -Country_Territory), by=c("neighbor" = "id")) %>% 
    select(EEZID = EEZID.x,Country_Territory,Neighbor_Territory) # select final column

    # Merge them with SAU data

  DBEM_Neighbors_Data <- EEZIDs_List %>% 
  left_join(EEZ_CellID, 
            by = "EEZID") %>% # Asign EEZ id to each country
  left_join(Neighbors, 
            by = "EEZID") %>% # Include all neighboring countries 
  filter(Neighbor_Territory != "NA") %>% 
  left_join(Coor,
            by = "INDEX") # Inclu

  return(DBEM_Neighbors_Data)

}

### Test function

# Subset data for testing function
# SAU_Regions <- as.data.frame(unique(World_EEZs$Name))

# Test_Countries <-c("Mexico (Atlantic)","Mexico (Pacific)","Guatemala (Caribbean)","Guatemala (Pacific)","Belize")
# Test <- World_EEZs[World_EEZs@data$Name %in% Test_Countries, ]

# Running function
# getNeighbors(Test)

#### Works beautifully!

@@

### Function `GetSppDist`

This function Read in species distribution from the observation data, ESM models and SAU data, still missing DBEM

```{r GetSppDist, eval = T, echo = T}

### Notes
# Function originally created by William (#WC)
# Juliano modifications (#JC)
# This funtion is used in estAbdBinTemp.R
# We are using only summer distribution

### Variables used
# spplisttemp <- Species list template. Information of the species used in the DBEM under TaxaDataC. 
# Distpath <- Path to where the initial distribution of the species is. AKA. Files that start with "S"

#### The function

GetSppDist=function(Spp,Model){

  Distpath <- "/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Distribution_Data/"
  INDEX <- seq(1,259200,1)

  # SAU Distributions
  if(Model == "SAU"){
    File_Name <- paste("SAU_Distribution/DIST_GOOD_SP_")
  }

  # Occurence
  if(Model == "Occurence"){
    File_Name <- paste("Occurence/OCCURENCE_JULIANO_")
  }

  # ENM Model
  if(Model == "ENM"){
    File_Name <- paste("ENM/ENM_JULIANO_")
  }

  # DBEM Data
  # if(Model == "DBEM"){
  #   Distpath <- "/Volumes/DATA/JULIANO_NEYMAR/PristineSeasData
  #   Final_Path <- "ENM"
  #   File_Name <- paste("ENM_JULIANO",Spp,".mat",sep="")
  # }

  if(Model == "All"){

    Models_List <- c(paste(Distpath,"SAU_Distribution/DIST_GOOD_SP_",Spp,".mat",sep =""),
                     paste(Distpath,"Occurence/OCCURENCE_JULIANO_",Spp,".mat",sep=""),
                     paste(Distpath,"ENM/ENM_JULIANO_",Spp,".mat",sep="")
    )

    # Jumps species not modeled. NOTE: We only need one as ENM, Occ and SAU Dis have all the same 939 spp
    if(file.exists(Models_List[1])){

    Load <- lapply(Models_List, FUN=R.matlab::readMat, na.strings=0)
    sppdist <- as.data.frame(bind_cols(Load))
    sppdist$INDEX <- INDEX
    colnames(sppdist) <- c("SAU","Occ","ENM","INDEX")

    getSppDist = sppdist

    }else{

      # print(paste("No info for this species"))
    }

  }else{

    # Merge paths
    SppPath <- paste(Distpath,File_Name,Spp,".mat",sep="")

    if(file.exists(SppPath) == TRUE){

    #Install (if needed) R.matlab package
    if(!require(R.matlab)){
      install.packages("R.matlab")
    }

    # Read Files

    sppdist <- as.data.frame(R.matlab::readMat(SppPath))
    colnames(sppdist) <- Model

    # Return 
    getSppDist=sppdist

    }else{
    print(paste("No info for this species",Spp, "in",Model))
  }

  }
}

#______ Test ______#

# Variables outside function
# head(getSppDist(600004,"SAU")) # Works
# head(getSppDist(600004,"Occurence")) # Works
# head(getSppDist(600004,"ENM")) # Works
# head(getSppDist(600004,"All")) # Works

@@

### Function `EstTransIndex`

@@

# Varibales needed
# Spp: Species to be analized, data with presence/absence per gridcell
# Index_EEZ: Reference list of all INDEX that fall within EEZs
# EEZ_Size: Number of grid-cells that each EEZ has
# Neighbors: Reference list of neighbouring countries

# Needs getSppDist

# The function
EstTransIndex <- function(SppDist,EEZ_Size,Neighbors){

  Index_EEZ <- Neighbors %>% 
    select(INDEX)
      # Result 1. Number of Countries that share the species
  # Trans_Index

  Trans_Spp <- SppDist %>%
    # gather("Model","Value",1:3) %>% 
    filter(
      INDEX %in% Index_EEZ$INDEX,
      # Value != 0
      SAU != 0
      ) %>%
    left_join(Neighbors,
              by = "INDEX") %>% # Match the species distribution to the Countries EEZs
    group_by(Name = Country_Territory,
             Neighbor_Territory#,
             # Model
             ) %>% 
    summarise(n_spp = n()) %>% # The number of cells present within each country's EEZ
    left_join(EEZ_Size,
              by = "Name") %>% 
    mutate(Trans_Index = n_spp/n_grid) %>%  # Include one transboundary species
    select(EEZID,everything())  # For the sake of my brain

@@

  Countries_with_Specie <- Trans_Spp %>% 
    filter(Trans_Index == 1) %>% 
    group_by(Neighbor_Territory) %>% 
    summarise()

@@

  EstTransIndex=Trans_Spp # A dataframe with each country and neighbouring states sharing the species

} # End function

### Testing Function ####
# SppDist <- Spp_Dist
# EEZ_Size <- EEZ_Size
# 
# Neighbors <- Neighbor_List

# Routine

```{r Control_Pannel, eval = T, echo = F}

#### TESTING PERIOD ####
# I'm substetting central america to do the whole analysis (smaller EEZs = faster computation) once I have figure out things I'll run for all

### Data 
# For now using only the subdata

## Spatial Test Data
# SAU_Regions <- as.data.frame(unique(World_EEZs$Name))

Test_Countries <-c("Guatemala (Caribbean)","Guatemala (Pacific)","Belize","Honduras","El Salvador","Costa Rica (Pacific)","Costa Rica (Caribbean)", "Panama (Caribbean)", "Panama (Pacific)")

Test <- World_EEZs[World_EEZs@data$Name %in% Test_Countries, ]

## Species Test Data

# DBEM exploited species list 
exploited_species_list <- fread("/Volumes/DATA/PROJECTION exploited species ENM/exploited_species_list.csv")

SAU_Data <- read.csv("/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Distribution_Data/SAU_Catch_Data_Test/SAU_Catch_Data_Test.csv")

#### Species per country

Spps <- SAU_Data %>% 
  filter(data_layer == "Reconstructed domestic catch") %>% 
  group_by(area_name,TaxonName = scientific_name) %>% 
  summarise(n()) %>% 
  left_join(exploited_species_list,
            by = "TaxonName")

# TakonKey <- Spps$TaxonKey[1:2]
TakonKey <- c("600004","600006",600107)
# Spp = 600988
TaxonKey <- unique(Spps$TaxonKey)

NOTA: Explorando resultados y hay que arrelgar la cuestion de los tres modelos

```{Routine, eval = T, echo = F}

# Step 1; EstNeighbordList 
Neighbor_List <- EstNeighbors(Test)
# head(Neighbor_List)

# Step 2; GetSppDist
# Spp_Dist <- GetSppDist(TakonKey[1],"All")
# head(Spp_Dist)

# Multiple TaxonKeys
Spp_Dist <- NULL
Spp_Dist <- lapply(TaxonKey, FUN = GetSppDist, "All")

Spp_Dist <- Spp_Dist[!sapply(Spp_Dist, is.null)]
# Step 3; EstTransIndex
# Spp_Trans <- EstTransIndex(Spp_Dist,EEZ_Size,Neighbor_List)
# Spp_Trans

# Multiple TaxonKeys
Spp_Trans <- NULL
Spp_Trans <- bind_rows(lapply(Spp_Dist, FUN=EstTransIndex, EEZ_Size = EEZ_Size, Neighbors = Neighbor_List))

Spp_Trans %>% 
  mutate(Uncert = ifelse(Trans_Index >= 0.10 & Trans_Index <= 0.25,"10 %",
                         ifelse(Trans_Index >= 0.25 & Trans_Index <= 0.50,"25 %",
                                ifelse(Trans_Index >= 0.50,">50 %","No Trans")))
         ) %>% 
  group_by(Name,Uncert) %>% 
    summarise(nspp=n()) %>% 
  ggplot(.,
         aes(
           x = Name,
           y = nspp
         )
         )+
  geom_bar(stat ="identity") +
  facet_wrap(~Uncert)
